Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Intro
Please check diff :)

Using the browser as a computing resource in volunteer computing presents several advantages, 
but it remains a challenge to fully harness the browser's capabilities and
to model the user's behavior so that those capabilities can be
leveraged optimally. In this paper, we present the results of several
evolutionary computation experiments with different implementations of
a volunteer computing framework called {\sf NodIO}, designed to be
easily deployable in using free cloud resources. We use  
%Pablo: "deployable in using free cloud resources" is correct?
% why not? - JJ
different implementations to find out which one is 
able to get the user to lend more computing cycles and test different
problems to check the influence it has on said performance.
% Pablo: so the objective is "to determine which browser conf. is able
% to get more computing power from the users", it isn't? 
% (to keep this in mind the rest of the paper). What is a "browser configuration"?
% (explain it with a phrase in the abstract)
% No, we don't. Please read the paper!!!! - JJ
% FERGU: Sorry, JJ, but the objective is not clear to me yet here, 
% even if the implementation comparison I wrote is in the conclusions as the first 
% objective of the paper. Also, the metrics used should be mentioned in the abstract.
% For example: "The first objective of this paper is to compare different implementations 
% of NodIO (with and without Web Workers), measuring the time spent by the users. 
% The second objective is to model the user behavior by comparing whatever".
% Also, the two compared implementations (w and w/ Web Worker) should be mentioned.
From these experiments we can already draw some conclusions, besides
the fact that volunteer computing can be a valuable computing
resource: the user has to be kept engaged in some way to be able to
use as many computing cycles as possible, the client has to be built
to use the computer capabilities fully, and, finally, the user
contributions follow a common statistical distribution. It is also
essential to be as open as possible with software and data.

@@

There is, then, space for other EC frameworks that do not make any of
these assumptions, such as {\sf NodIO} presented in
\cite{2016arXiv160101607Manom}. {\sf NodIO} implements the server of
client/server application and has some support to program an
evolutionary algorithm embedded in pages served by them which can then
communicate via that server. The client part can run in any
web-enabled device, including many over which developers have no
control, that is, volunteers.

@@

%PABLO: He añadido el siguiente párrafo para el issue #31
There exist other volunteer computing frameworks written in compiled languages,
such as C++ or Java, that, in principle, can achieve better performance than
JavaScript. However, one of the advantages of our architecture is that
clients do not need to install anything on their machines, they just
%Pablo: According English Stack Exchange, you install "ON" a machine (but "IN" a directory)
need to open a web page  
in a compatible, that is, almost any, browser, such as Chrome, Safari or Firefox, which are
available in almost every desktop or mobile device. Another advantage is
that the source code can be directly examined by volunteers, because it is not
compiled as other options such as Java Applets, ActionScript or
ActiveX are. Also, as we will discuss in this paper, it can provide some extra
features such as enhanced security or operating system
independence. This system, however, is not intended to substitute
frameworks running on dedicated hardware or the cloud, but to
%Pablo: same here, I learnt this with my paper of EAs "on clusters"
complement it. In general, several volunteers will be needed to
overcome the overhead involved in using a client-server architecture
and this number of volunteers is obviously not guaranteed.

@@

One of the aspects that must be taken into account in volunteer based
frameworks is its {\em social} aspect in the design, with 
issues related to security, trust, and privacy among others. 
The computing system becomes then a {\em techno-social system}
\cite{vespignani2009predicting}. In that sense, its social aspects
cannot be separated from its technical specifications, with one
feeding into the other. One of these aspects is safety for the user
and the scientist running it: the user must be reasonably sure that the
experiment is not misusing its resources and the scientist must have a
certain safety in the sense that its server is not going to be
attacked via cheating or denial or service attacks. We have decided to
approach this in a pragmatic way. Including many security features
would decrease performance; on the other hand, we publish all data
resulting from the experiment once private information, such as IPs,
has been processed to a untraceable hash. The experiments themselves
are a proof that there are enough users trusting the experiment to
give it its CPU time and that there has been no attempt to cheat by
sending specially crafted chromosomes or to attack the server to make
it stop working because there was no incentive in doing so. The
system, besides, is running in a dedicated Platform as a Service
container or Virtual Private Server so accessing the server would only
show a few data obtained from the experiment, nothing more. Making it
free software ensured this was known by anyone in the experiment, via
links in the web page that hosted it.

In this paper, we will carry out three experiments, after establishing
the baseline of the performance that can be expected in a JavaScript
based system and its comparison with other languages in
\cite{2016arXiv160101607Manom}. First, several experiments using a
Platform as a Service, and a simple Ajax client that does not
restart. Second, we change a bit the client so that it uses web
workers, that is, it runs in the background and it restarts. Third, we use the same
% Pablo: it restarts when the solution is found (it has not been mentioned yet)
client, but with a harder problem that cannot be solved by a single
volunteer and thus needs the contribution of many volunteers to
proceed. With these three experiments we want mainly to try and model
the behavior of users and how different problems and implementations
have an influence on it. 
%Pablo: To do this we compare the time spent by each user because...

So, the rest of the paper is organized as follows: next we present the
state of the art in web-based distributed
computational systems along with attempts to predict and model its
behavior  in Section \ref{sec:soa}. A brief presentation of the
methodology used and the results of the experiments are presented in
%Pablo: mention the problems used in the first experiments
Section \ref{sec:exp1}, including new results using web workers on the
same problem and on a new problem, Rastrigin or F15 function. 
%Pablo: Rastrigin is not the owner of the function
Finally, conclusions and future lines of work are presented in Section
\ref{sec:conclusion}.

>>>>

Using the browser as a computing resource in volunteer computing presents several advantages, 
but it remains a challenge to fully harness the browser's capabilities and
to model the user's behavior so that those capabilities can be
leveraged optimally. In this paper, we present the results of several
evolutionary computation experiments with different implementations of
a volunteer computing framework called {\sf NodIO}, designed to be
easily deployable on freely available cloud resources. We use  
%Pablo: "deployable in using free cloud resources" is correct?
% why not? - JJ
% is easy deployable? yes! where? in free cloud resources! 
% is easy deployable? yes! when?  (when) using free cloud resources
% is one or the other, I like "on"      
different implementations to find out which one is 
able to get the user to lend more computing cycles and test different
problems to check the influence it has on said performance.
% Pablo: so the objective is "to determine which browser conf. is able
% to get more computing power from the users", it isn't? 
% (to keep this in mind the rest of the paper). What is a "browser configuration"?
% (explain it with a phrase in the abstract)
% No, we don't. Please read the paper!!!! - JJ
% FERGU: Sorry, JJ, but the objective is not clear to me yet here, 
% even if the implementation comparison I wrote is in the conclusions as the first 
% objective of the paper. Also, the metrics used should be mentioned in the abstract.
% For example: "The first objective of this paper is to compare different implementations 
% of NodIO (with and without Web Workers), measuring the time spent by the users. 
% The second objective is to model the user behavior by comparing whatever".
% Also, the two compared implementations (w and w/ Web Worker) should be mentioned.
From these experiments we can already draw some conclusions, besides
the fact that volunteer computing can be a valuable computing
resource: the user has to be kept engaged in some way to be able to
use as many computing cycles as possible, the client has to be built
to use the computer capabilities fully, and, finally, the user
contributions follow a common statistical distribution. It is also
essential to be as open as possible with software and data.

@@

There is, then, space for other EC frameworks that do not make any of
these assumptions, such as {\sf NodIO} presented in
\cite{2016arXiv160101607Manom}. {\sf NodIO} implements the server of a
client/server application and has some support to program an
evolutionary algorithm embedded in pages served by them which can then
communicate each other via that server. The client part can run in any
web-enabled device, including many over which developers have no
control, that is, volunteers.

@@

%PABLO: He añadido el siguiente párrafo para el issue #31
There exist other volunteer computing frameworks written in compiled languages,
such as C++ or Java, that, in principle, can achieve better performance than
JavaScript. However, one of the advantages of our architecture is that
clients do not need to install anything on their machines, they just
%Pablo: According English Stack Exchange, you install "ON" a machine (but "IN" a directory)
need to open a web page  
in a compatible, that is, almost any, browser, such as Chrome, Safari or  Firefox, which are
available in almost every desktop or mobile device. Another advantage is
that the source code can be directly examined by volunteers, because it is not
compiled as other options such as Java Applets, ActionScript or
ActiveX are. Also, as we will discuss in this paper, it can provide some extra
features such as enhanced security or operating system
independence. This system, however, is not intended to substitute
frameworks running on dedicated hardware or the cloud, but to
%Pablo: same here, I learnt this with my paper of EAs "on clusters"
complement them. In general, several volunteers will be needed to
overcome the overhead involved in using a client-server architecture
and this number of volunteers is obviously not guaranteed.

@@

%Mario: Rewrote: 
%One of the aspects that must be taken into account in volunteer based
%frameworks is its {\em social} aspect in the design, with 
The {\em social} aspect must be taken into account when designing 
volunteer based frameworks, with 
issues related to security, trust, and privacy among others. 
The computing system becomes then a {\em techno-social system}
\cite{vespignani2009predicting}. In that sense, its social aspects
cannot be separated from its technical specifications, with one
feeding into the other. One of these aspects is safety for the user
and the scientist running it: the user must be reasonably sure that the
experiment is not misusing its resources and the scientist must have a
certain safety in the sense that its server is not going to be
attacked via cheating or denial of service attacks. We have decided to
approach this in a pragmatic way. Including many security features
would decrease performance; on the other hand, we publish all data
resulting from the experiment once private information, such as IPs,
has been processed to an untraceable hash. The experiments themselves
are a proof that there are enough users trusting the experiment to
give it its CPU time and that there has been no attempt to cheat by
sending specially crafted chromosomes or to attack the server to make
it stop working because there was no incentive in doing so. The
system, besides, is running in a dedicated Platform as a Service (PaaS)
container or Virtual Private Server so accessing the server would only
show the few data obtained from the experiment, nothing more. Making it
free software ensured this was known by anyone in the experiment, via
links in the web page that hosted it.

In this paper, we will carry out three experiments, after establishing
the baseline of the performance that can be expected in a JavaScript
based system and its comparison with other languages in
\cite{2016arXiv160101607Manom}. First, several experiments are performed using a PaaS, and a simple Ajax client that does not
restart when a solution is found. Second, we change the client a bit so that it uses web workers, that is, it runs in the background process and it restarts. Third, we use the same
% Pablo: it restarts when the solution is found (it has not been mentioned yet)
client, but with a harder problem that cannot be easily solved by a single
volunteer and thus needs the contribution of many volunteers to
proceed. 
%Mario: refrased: 
%With these three experiments we want mainly to try and model
With these three experiments our main objective is to try and model
the behavior of users and learn how different problems and implementation choices have an influence on it.

%Pablo: To do this we compare the time spent by each user because...

So, the rest of the paper is organized as follows: next we present the
state of the art in web-based distributed
computational systems along with attempts to predict and model their
behavior in Section \ref{sec:soa}. A brief presentation of the
methodology used and the results of the experiments are presented in
%Pablo: mention the problems used in the first experiments
Section \ref{sec:exp1}, including new results using web workers on the
same problem and on a new problem, a modified Rastrigin (F15) function. 
%Pablo: Rastrigin is not the owner of the function
%Mario: In many places is called  Rastrigin's function because it was proposed by him.
Finally, conclusions and future lines of work are presented in Section
\ref{sec:conclusion}.
